---
output:
  github_document
title: "DesignLibrary: A Library of Common Research Designs"
always_allow_html: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

[![Coverage Status](https://coveralls.io/repos/github/DeclareDesign/DesignLibrary/badge.svg?branch=master)](https://coveralls.io/github/DeclareDesign/DesignLibrary?branch=master)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/DesignLibrary)](https://cran.r-project.org/package=DesignLibrary)
[![minimal R version](https://img.shields.io/badge/R%3E%3D-3.4.0-6666ff.svg)](https://cran.r-project.org/)
[![packageversion](https://img.shields.io/badge/Package%20version-0.1.5.9999-orange.svg?style=flat-square)](commits/master)

**DesignLibrary** provides simple interface to build designs using the package **DeclareDesign**. In one line of code users can specify the parameters of individual designs and diagnose their properties. The designers can also be used to compare performance of a given design across a range of combinations of parameters, such as effect size, sample size, assignment probabilities and more.

---

## Designs

```{r, echo=FALSE, message=FALSE}
library(fontawesome)
library(knitr)
library(tidyverse)
library(kableExtra)

add_link <- function(url, fa_string, tooltip) {
  paste0(
    "<a data-toggle = 'tooltip' title = '",
    tooltip,
    "' href=\"",
    url,
    "\">",
    as.character(fa(fa_string, height = "15px", fill = "#2860F6")),
    "</a>"
  )
}

overview <- read_csv("inst/extdata/overview.csv")

overview <-
  overview %>%
  mutate(Vignette = add_link(url = paste0("https://declaredesign.org/r/designlibrary/articles/", vignette, ".html"), 
                                  fa_string = "fab fa-readme", 
                                  "Read description of design"),
         Designer = add_link(url = paste0("https://declaredesign.org/r/designlibrary/reference/", designer, ".html"), 
                                  fa_string = "fas fa-pencil-alt", 
                                  "Open designer documentation"),
         `Design Inspector` = add_link(url = paste0("https://eos.wzb.eu/ipi/DDinspector/?import_library=", design), 
                                       fa_string = "fas fa-info-circle", 
                                       "Open in DeclareDesign Inspector"),
         design_title = str_replace_all(design, "_", " "),
         design_title = str_to_title(design_title),
         Design = str_replace(design_title, "Iv", "IV"))


overview %>%
  select(Design, Vignette, Designer, `Design Inspector`) %>%
  kable(escape = F, align = c("l", "c", "c", "c")) %>%
  kable_styling()
```


## Installing the design library

To install the latest stable release of **DesignLibrary**, please ensure that you are running version 3.4 or later of R and run the following code:

```{r, eval=F}
install.packages("DesignLibrary")
```

If you would like to use the latest development release of **DesignLibrary**, please ensure that you are running version 3.4 or later of R and run the following code:

```{r, eval=F}
devtools::install_github("DeclareDesign/DesignLibrary", keep_source = TRUE)
```

## Contributing designs and designers

We welcome contributions to the library! 

- You can [submit static designs](https://declaredesign.org/r/designlibrary/articles/how_to_write_and_contribute_designs.html) made in `DeclareDesign`, which will live as properly attributed entries in the library on our website
- Or you can [submit designer functions that generate designs](https://declaredesign.org/r/designlibrary/articles/how_to_write_and_contribute_designers.html), which may be added to the CRAN version of the package

---

This project is generously supported by a grant from the [Laura and John Arnold Foundation](http://www.arnoldfoundation.org) and seed funding from [Evidence in Governance and Politics (EGAP)](http://egap.org).
