---
output:
  github_document
title: "DesignLibrary: A Library of Common Research Designs"
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

[![Travis-CI Build Status](https://travis-ci.org/DeclareDesign/DesignLibrary.svg?branch=master)](https://travis-ci.org/DeclareDesign/DesignLibrary)
[![Coverage Status](https://coveralls.io/repos/github/DeclareDesign/DesignLibrary/badge.svg?branch=master)](https://coveralls.io/github/DeclareDesign/DesignLibrary?branch=master)

**DesignLibrary** provides a library of functions that return designs built using the package **DeclareDesign**. In one line of code users can specify the parameters of individual designs and diagnose their properties. The designers can also be used to compare performance of a given design across a range of combinations of parameters, such as effect size, sample size, assignment probabilities and more.

[Check out the online version of the library here](https://declaredesign.org/library/).

---

## Installing the design library

To install the latest stable release of **DesignLibrary**, please ensure that you are running version 3.4 or later of R and run the following code:

```{r, eval=F}
install.packages("DesignLibrary")
```

If you would like to use the latest development release of **DesignLibrary**, please ensure that you are running version 3.4 or later of R and run the following code:

```{r, eval=F}
devtools::install_github("DeclareDesign/DesignLibrary", keep_source = TRUE)
```

---

This project is generously supported by a grant from the [Laura and John Arnold Foundation](http://www.arnoldfoundation.org) and seed funding from [Evidence in Governance and Politics (EGAP)](http://egap.org).
