---
title: "Example Design"
output: rmarkdown::html_vignette
bibliography: bib.bib
vignette: >
  %\VignetteIndexEntry{Example Design}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

You should make sure to install any required packages. The example design uses `systemfit` as well as `DeclareDesign`.

```{r eval = FALSE}
install.packages("systemfit")
install.packages("DeclareDesign")
install.packages("knitr")
install.packages("RCurl")
```

Load the packages:

```{r,message=FALSE, eval = FALSE}
library(systemfit)
library(DeclareDesign)
library(knitr)
library(RCurl)
```

Here is one way to load external data for use in the design:

```{r,warning = FALSE, eval = FALSE}
online_data <- read.csv(file = "https://declaredesign.org/example_data.csv",
                        header = TRUE)

rho <- with(online_data, rho)
tau_A <- with(online_data, tau_A)
tau_B <- with(online_data, tau_B)
```

## Design Declaration

You may want to give an overview of the design, and provide any background info.

- **M**odel: 
    
    We have two untreated potential outcomes, $Y_A(0)$ and $Y_B(0)$ that are distributed standard normally and correlated with each other at $\rho = .2$. We have $Y_A(1) = Y_A(0) + \tau_A$ and  $Y_B(1) = Y_B(0) + \tau_B$. 
  
- **I**nquiry: 
    
    We want to know the value of $E[Y_A(1) - Y_A(0)]$. 

- **D**ata strategy: 
    
    Half the units are assigned to a treatment, $Z_A$, that only affects $Y_A$. Units are also independently assigned to $Z_B$, a treatment that only affects $Y_B$. 

- **A**nswer strategy: 
    
    We compare two strategies: in one we simply regress $Y_A$ on an indicator for $Z_A$, and in the other we use include in our system of seemingly unrelated regressions (SUR) a second equation for $Y_B$ on $Z_B$. 

## Design code

Very often, designs require custom functions and data cleaning. You may want to do that separately from your design declaration. Here we make a custom `sur()` function to use the seemingly unrelated regression estimator in `systemfit::systemfit`. 

```{r, eval = FALSE}
sur <- function(data){
  fit <- data.frame(summary(systemfit(formula = list(YA ~ ZA, 
                                                     YB ~ ZB),
                                      data = data,
                                      method = "SUR"))$coefficients)
  names(fit) <- c("estimate","std.error","statistic","p.value")
  fit$conf.low <- with(fit, estimate - 1.96 * std.error)
  fit$conf.high <- with(fit, estimate + 1.96 * std.error)
  fit <- cbind(estimator_label = "SUR",
               term = row.names(fit),
               fit, 
               df = NA, 
               outcome = "YA",
               estimand_label = "ate_YA")
  row.names(fit) <- NULL
  fit <- subset(fit, term == "eq1_ZA")
  return(fit)
}
```

Now we declare the design in code.

```{r, eval = FALSE}
my_design <- 
  # Model
  declare_population(N = 100,
                     YA_0 = rnorm(N),
                     YB_0 = rnorm(N, YA_0 * rho, sqrt(1 - rho^2))) +
  declare_potential_outcomes(YA ~ YA_0 + ZA * tau_A,
                             assignment_variable = "ZA") +
  declare_potential_outcomes(YB ~ YB_0 + ZB * tau_B,
                             assignment_variable = "ZB") +
  # Inquiry
  declare_estimand(ate_YA = mean(YA_ZA_1 - YA_ZA_0)) +
  # Data Strategy
  declare_assignment(m = 50, assignment_variable = "ZA") +
  declare_assignment(m = 50, assignment_variable = "ZB") +
  declare_reveal(YA,ZA) +
  declare_reveal(YB,ZB) +
  # Answer Strategy
  declare_estimator(YA ~ ZA, model = lm_robust, 
                    estimand = "ate_YA", label = "lm_robust") +
  declare_estimator(handler = sur) 
```


## Design diagnosis

You may want to display features of the design here.

```{r, eval = FALSE}
my_design_cor <- redesign(design = my_design, rho = .8)
diagnosis <- diagnose_design(my_design, my_design_cor, sims = 25)
kable(reshape_diagnosis(diagnosis), digits = 2)
```

We see that the SUR approach exhibits much lower variance in estimates and much higher power when $\rho$ is high.

