---
title: "How to Write and Contribute Designs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to Write and Contribute Designs}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DesignLibrary)
```

This guide provides instructions for contributing static design objects to the library at [DeclareDesign.org/library](DeclareDesign.org/library). Please note, designs should be submitted as `.Rmd` files (vignettes), and will not be added to the CRAN version of the `DesignLibrary`. 

First, install `DeclareDesign` in `R`:

```{r,eval=FALSE}
install.packages("DeclareDesign")
```

A design is an object of class `design`, created by concatenating a series of `declare_*` calls using the `+` operator in `DeclareDesign`. 

## Contributing designs

Your design should be submitted to `DesignLibrary` as an `.Rmd` vignette located in the vignette subfolder of the github repository: [see here](https://github.com/DeclareDesign/DesignLibrary/tree/master/vignettes). 


You should submit your design via a pull request to [https://github.com/DeclareDesign/DesignLibrary/](https://github.com/DeclareDesign/DesignLibrary/). This will typically involve [forking the repository](https://guides.github.com/activities/forking/), creating a new branch, adding your design to the vignettes folder, and [then submitting a pull request](https://guides.github.com/activities/forking/) to the `DeclareDesign/DesignLibrary` repository. 

When you submit the pull request, you should see a pull request template, which contains a checklist. This checklist covers additional points addressed below.

## Meta-data

In order for your design to appear on the table at [DeclareDesign.org/library](DeclareDesign.org/library), you will need to add some meta-data about your design to the `overview.csv` file located at [DesignLibrary/inst/extdata](https://github.com/DeclareDesign/DesignLibrary/tree/master/inst/extdata/overview.csv).

You should add a row with: 

  - the name of your design under `design`
  - leave `designer` blank
  - put the name of the vignette file without the `.Rmd` extension under `vignette`
  - add any keywords under `keywords`, separated by commas
  - leave `shiny` blank
  - you can put your name under `author`
  - and a link to your website or twitter handle under `author_url`

## External data

Very often, more complicated designs will rely on external data. If your design requires external data to run, please make sure that it is sourced into your vignette through a publicly accessible URL.  We will not accept any objects in a pull request aside from an `.Rmd` file.

## If in doubt...

Just [post an issue](https://github.com/DeclareDesign/DesignLibrary/issues) and we'll get right on it.





