---
title: "Mediation Analysis Design"
output: rmarkdown::html_vignette
bibliography: bib.bib
nocite: |
  @Gerber2012
  @imai2011
designer: "mediation_analysis_designer.Rd"
vignette: >
  %\VignetteIndexEntry{Mediation Analysis Design}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DesignLibrary)
library(knitr)
```

By randomly assigning units to treatment we can determine whether a treatment affects an outcome but not *why* or *how* it might affect it. Identifying causal mechanisms is not a simple task. It involves complex potential outcomes and a mediating variable that is generally not assigned at random  and is not a pre-treatment covariate (since it's affected by the treatment). Researchers often use regression-based approaches to identify causal mechanisms but these rely on assumptions that sometimes can't be met.

For this analysis we assume that there is a non-zero average treatment effect (ATE) of $Z$ on $Y$. Our main interest lies in decomposing the ATE into *direct* and *indirect effects*. The indirect effect is channeled from the treatment $Z$ to the outcome $Y$ through a mediator $M$ and the direct effect runs directly from $Z$ to $Y$.

## Design Declaration

- **M**odel:  
    
    We specify a population of size $N$. Individuals from this population have two potential outcomes related to the mediator.

    * $M_i(Z_i=0):$ The value for the mediator $M$ when the unit $i$ is in the control group, and
    
    * $M_i(Z_i=1):$ the value for the mediator $M$ when the unit $i$ is treated.

    Additionally, individuals have four potential outcomes related to $Y$.

    Two can be observed under treatment or control conditions.

    * $Y_i(Z_i=0 , M_i(Z_i=0)):$  the outcome of unit $i$ when the treatment is absent and the mediator takes the value it would when the treatment is absent.

    * $Y_i(Z_i=1 , M_i(Z_i=1)):$ the outcome of unit $i$ when the treatment is present and the mediator takes the value it would when the treatment is present.

    And two complex potential outcomes.

    * $Y_i(Z_i=1 , M_i(Z_i=0)):$ the outcome of unit $i$ when the treatment is present *but* the mediator takes the value it would when the treatment is *absent*.

    * $Y_i(Z_i=0 , M_i(Z_i=1)):$ the outcome of unit $i$ when the treatment is absent *but* the mediator takes the value it would when the treatment is *present*.

Thus the data generating process we specify defines $Y$ as a function of $M$ and $Z$ and $M$ as a function of $Z$.
  
- **I**nquiry: 

    We are interested in the average effects of the treatment on the mediator, $M$, the direct average effects of the treatment on $Y$ and the effects on $Y$ from $Z$ that run through $M$.

- **D**ata strategy:  

    We use assign units to treatment using complete random assignment.

- **A**nswer strategy:  

    First, we regress $M$ on $Z$. Then we regress $Y$ on $M$ and $Z$. 


```{r, code = get_design_code( mediation_analysis_designer()), eval=TRUE}
```

## Takeaways

We diagnose two versions of this design: one in which the correlation between the error term of the mediator regression and one of the outcome regression ($\rho$) is greater than zero, and another in which $\rho$ equals zero.


```{r,eval = TRUE}
designs <- expand_design(mediation_analysis_designer, rho = c(0,.5))
diagnosis <- diagnose_design(designs)
```

```{r,eval = TRUE, echo = FALSE}
kable(reshape_diagnosis(diagnosis)[,-1], digits = 2)
```

Our diagnosis indicates that when the error terms are not correlated, the direct and indirect effects can be estimated without bias. By contrast, when $\rho$ does not equal zero, the regression underestimates the effect of the mediator on $Y$ and overstates the direct effects of $Z$ on $Y$.

Unfortunately, the assumption of no correlation is not always guaranteed, since $M$ is not assigned at random and might be correlated with $Y$.

