---
title: "Mediation Analysis Design"
output: rmarkdown::html_vignette
bibliography: bib.bib
nocite: |
  @Gerber2012
  @imai2011
designer: "mediation_analysis_designer.Rd"
vignette: >
  %\VignetteIndexEntry{Mediation Analysis Design}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DesignLibrary)
library(knitr)
```

By randomly assigning units to treatment we can determine whether a treatment affects an outcome but not *why* or *how* it might affect it. Identifying causal mechanisms, which is arguably one of the most useful findings in science, requires a treatment with nonzero effects and the specification of one mediator.  


For this analysis we assume that the average treatment effect (ATE) of $Z$ on $Y$ is statistically different from zero.  Our main interest lies in decomposing the ATE into *direct* and *indirect effects*.  The indirect effect is channeled from the treatment $Z$ to the outcome $Y$ through a mediator $M$ and the direct effect runs directly from $Z$ to $Y$.


Identifyng causal mechanisms is not a simple task. It involves complex potential outcomes (more on this below) and a mediating variable that was not assigned at random  and is not a pre-treatment covariate (it's affected by the treatment!). Researchers often use regression based approaches to identify causal mechanisms but these rely on assumptions that sometimes can't be met.

We define below a mediation analysis design.

### Design Declaration

- **M**odel:  
We specify a population of size $N$. Individuals from this population have two potential outcomes related to the mediatior.

* $M_i(Z_i=0):$ The value for the mediator $M$ when the unit $i$ is in the control group, and
* $M_i(Z_i=1):$ the value for the mediator $M$ when the unit $i$ is treated.

Additionally, individuals have four potential outcomes related to $Y$.

Two potential outcomes that can be observed under treatment or control conditions.

  * $Y_i( Z_i=0 , M_i(Z_i=0)):$  the outcome of interest given that the unit  $i$ is in the control group and the mediator that would be observed if the unit $i$ is in the control group, and

* $Y_i( Z_i=1 , M_i(Z_i=1)):$ the outcome of interest given that the unit  $i$ is in the treatment group and the mediator that would be observed when the unit  $i$ is in the control group.

And two complex potential outcomes.

* $Y_i( Z_i=1 , M_i(Z_i=0)):$ the outcome of interest given that unit $i$ is treated and the value of the mediator hat would be observed in the control condition.

* $Y_i( Z_i=0 , M_i(Z_i=1)):$ the outcome of interest given that unit $i$ remains untreated and the value of the mediatorhat would be observed in the treatment condition


  
- **I**nquiry: We are interested in the average effects of the treatment on the mediatior $a$, the direct average effects of the treatment on $Y$ and the effects $Z$ that run through $M$.

- **D**ata strategy:  We use assign units to treatment using complete random assigment.

- **A**nswer strategy:  First, we regress $M$ on $Z$. Then we regress $Y$ on $M$ and $Z$. Our estimators are the coefficient of regressors.

In code:


```{r, code = get_design_code( mediation_analysis_designer()), eval=TRUE}

```

### Diagnosis

Let us first diagnose this design when the correlation between the error term of the mediator regression and the error term of the outcome regression.
```{r,eval = TRUE}
diagnosis <- diagnose_design( mediation_analysis_designer(rho = 0.5))
```


```{r,eval = TRUE, echo = FALSE}
kable(reshape_diagnosis(diagnosis)[, -c(1, 4:6, 13,14)], digits = 2)
```



Now, let us see what happens with the bias when the correlation between the error terms is zero.

```{r,eval = TRUE}
diagnosis <- diagnose_design( mediation_analysis_designer(rho = 0))
```

```{r,eval = TRUE, echo = FALSE}
kable(reshape_diagnosis(diagnosis)[, -c(1, 4:6, 13,14)], digits = 2)
```

### Further Reading


