---
title: "ResearchDesigns: a library of declared designs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ResearchDesigns: a library of declared designs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 4
)
# Prefer source designs when developing from the package tree / Dropbox checkout
if (dir.exists("../inst/designs")) {
  options(ResearchDesigns.root = normalizePath("..", winslash = "/"))
}
library(ResearchDesigns)
library(DeclareDesign)
library(ggplot2)
```

# The idea

To contribute a design to **ResearchDesigns** you just have to place a self-contained `.R` file in `inst/designs/`. That file declares a design and, optionally, a short YAML header. Once it is there, the design can be loaded, modified, diagnosed, and browsed in the bundled Shiny app.

For the contribution workflow (fork, add file, refresh, check, pull request), see the vignette *Contributing a design* or the Shiny **Contribute** tab.


# What a design file looks like

Here is the starter design `two_arm_simple`.

<details>
<summary><code>inst/designs/two_arm_simple.R</code></summary>

```{r, echo = FALSE, results = "asis"}
path <- system.file("designs", "two_arm_simple.R", package = "ResearchDesigns")
if (!nzchar(path) || !file.exists(path)) {
  root <- tryCatch(
    ResearchDesigns:::find_package_root(),
    error = function(e) NULL
  )
  if (!is.null(root)) {
    path <- file.path(root, "inst", "designs", "two_arm_simple.R")
  }
}
code <- if (file.exists(path)) {
  paste(readLines(path, warn = FALSE), collapse = "\n")
} else {
  "# two_arm_simple.R not found in this build"
}
cat("```r\n", code, "\n```\n", sep = "")
```

</details>

The header is optional. At minimum you need a design object (by default named `design`). Parameter tips are plain strings under `params:`; quote keys such as `"N"` and `"b"`.

As soon as that file sits in `inst/designs/`, it is available to `make_design()`, `get_args()`, `get_code()`, the maintainer audits, and `run_shiny()`.

# What designs are included?

```{r list-designs}
list_designs()
```

Call `list_designs()` to see ids and labels. Use `design_info("two_arm_simple")` or `get_args("two_arm_simple")` for one design, or `as.data.frame(list_designs(discover_params = TRUE))` for the table including redesignable parameters.

# Use ResearchDesigns to make a design

```{r make-default}
my_design <- make_design("two_arm_simple")
my_design
```


# Change parameters and simulate

Parameters are taken from the design itself. Pass new values to `make_design()`; under the hood this uses DeclareDesign's `redesign()`. A core set of designs also keep DesignLibrary names, so `two_arm_designer(N = 40, ate = 0.2)` is the same idea as `make_design("two_arm_flexible", N = 40, ate = 0.2)`.

```{r make-and-simulate}
make_design("two_arm_simple", b = 0.3) |>
  simulate_design(sims = 2)
```

You can now change all parameters in one go.

```{r make-and-simulate-grid}
make_design("two_arm_simple", b = c(0, .3, .6), N = c(20, 1000)) |>
  simulate_design(sims = 100) |>
  ggplot(aes(estimate, p.value, color = factor(b))) + 
  geom_point()  + facet_grid(N~.)
```

# Browse in Shiny

```{r shiny, eval = FALSE}
run_shiny()
```

The app opens on a searchable library table. Click a row to open the design, expand its profile, run it once, diagnose it, or redesign parameters. For a server deploy: install the package, run `install_library_dependencies()`, then `copy_library_shiny(dest)`.

# Maintainer tools

A short set of helpers keeps the library honest:

- `contributor_checklist()` — what a design file should satisfy
- `audit_designs()` — load each design and check that YAML `params:` names match redesignable objects
- `bake_previews()` — write compact diagnosis previews (`sims = 100` by default) into `inst/previews/`
- `refresh_library()` — index, audit, and bake in one step

Run maintainer helpers from the package source tree, or set:

```{r maintainer-root, eval = FALSE}
options(ResearchDesigns.root = "C:/path/to/ResearchDesigns")
refresh_library()
```

That is the whole loop: drop in a design file, use it from R or Shiny, and refresh the library when you change the set. Step-by-step contribution guidance is in `vignette("contributing", package = "ResearchDesigns")`.
